Mallows, Black-box combinatorial optimization, limited budget, ???

### 2020-10-24 paper/main.tex
improve 5

@@

\newcommand{\Prob}{\ensuremath{p}}
\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}

@@

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model for Black-Box Optimization}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $\minit$: number of initial permutations, $m$: total budget, $\rho$: learning rate
    %\\ ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \State $S \assign $ generate $\minit$ permutations uniformly at random
    \State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
    \For{$i \assign \minit$ \textbf{to} $m$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}
    \State linear decrease on $\theta$
    \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    \State $S \assign S \cup \{\sigma\}$, $\quad F \assign F \cup \{f(\sigma)\}$
    \EndFor 
    \State \Return $S$, $F$
  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
  %   \State linear decrease on $\theta$
  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
  %   \State $S \assign S \cup \{\sigma\}$
  %   \State $F \assign F \cup \{f(\sigma)\}$
  %   \EndFor 
  %   \State \Return $S$, $F$
 \end{algorithmic}
\end{algorithm}

As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated. Then, using the current sample $S$, the uBorda algorithm is applied to find a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}). Together with the parameter $\theta$, they define a Mallows model (Eq.~\ref{eq:MM}) from which we sample a new candidate permutation (line~\ref{line:sample}). This candidate permutation (solution) is evaluated and added to the sample $S$. The process is repeated up to a given maximum budget of evaluations $m$.\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it. Can we give a bit more detail?}

@@

\subsubsection{CEGO.}
We use the implementation of GECO provided by the
authors.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
Following the authors of
CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
the surrogate models, with population size of 20, crossover rate of 0.5,
mutation rate of 1, tournament selection of size 2 and probability of 0.9,
interchange mutation (i.e., exchanging two randomly selected elements) and
cycle crossover for permutations. To generate each new solution, CEGO runs the
GA using the model as a surrogate of the true fitness function.  The authors of
CEGO used a budget of $10^4$ evaluations of the surrogate-model for each run of
the GA. However, with such a setting, one run of CEGO with a budget of $m=400$
evaluations of the true fitness function requires more than a week of
computation time, which was infeasible for our study. Hence, we set the budget of the GA to $10^3$. Although this means that the results of CEGO reported here could, in principle, improved with a higher GA budget  we believe

Although it is never stated in the original paper, the implementation of CEGO
generates a set of initial solutions of size $\minit=10$ by means of a
max-min-distance sequential design: new solutions are added to the set
sequentially by maximizing the minimum distance to solutions already in the
set. These initial solutions are then evaluated on the actual objective
function and the result is used to build the initial surrogate
model.% \citet{ZaeStoBar2014:ppsn} uses 10

The CPU time required for a single run of CEGO using these settings is larger
than a week.  For simplicity, we use Kendall's $\tau$ distance~(Eq.~\ref{eq:kendall_decomp}) in all
experiments\MANUEL{\citet{ZaeStoBar2014:ppsn} claims this is called Swap
  distance, could you double-check?}. Nevertheless,
\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
when using Hamming distance. We plan to extend our analysis of UMM to other
distance measures and to the dynamic selection of distance
measures~\citep{ZaeStoBar2014:ppsn}.

In all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
the actual objective function. In a white-box context, state-of-the-art
algorithms for the problems considered here typically evaluate
\textcolor{red}{hundreds? thousands?} of solutions, thus, the budget considered
here for the black-box context is extremely limited.

@@

For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$?}

>>>>

\newcommand{\Prob}{\ensuremath{p}}
\newcommand{\minit}{\ensuremath{m_\text{ini}}\xspace}
%\newcommand{\FEmax}{\ensuremath{\textit{FE}_{\max}}}
\newcommand{\FEmax}{\ensuremath{m}}

@@

@@

\begin{algorithm}[t]
 \caption{UMM: Unbalanced Mallows Model for Black-Box Optimization}
 \label{alg:umm}
  \begin{algorithmic}[1]
    \Require $\minit$: number of initial permutations, $\FEmax$: total budget, $\rho$: learning rate
    %\\ ?, $\FEmax$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
    \State $S \assign $ generate $\minit$ permutations uniformly at random
    \State $F \assign \{ f(\sigma) \mid \sigma \in S\}$ \Comment{Evaluate them}
    \For{$i \assign \minit$ \textbf{to} $\FEmax$ \text{evaluations}}
    \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$\label{line:uborda}
    \State linear decrease on $\theta$
    \State $\sigma \assign$ sampled from  $MM(\hat\sigma_0, \theta)$\label{line:sample}
    \State $S \assign S \cup \{\sigma\}$, $\quad F \assign F \cup \{f(\sigma)\}$
    \EndFor 
    \State \Return $S$, $F$
  %   \Require $n$: ?, $m$: budget of evaluations,\newline $S$: initial permutations, $F$: their fitness values
  %   \For{$i \assign 1$ \textbf{to} $m$ \text{evaluations}}
  %   \State $\hat\sigma_0 \assign \text{uBorda}(S, F,\rho)$
  %   \State linear decrease on $\theta$
  %   \State $\sigma \assign$ sample from a $MM(\hat\sigma_0, \theta)$
  %   \State $S \assign S \cup \{\sigma\}$
  %   \State $F \assign F \cup \{f(\sigma)\}$
  %   \EndFor 
  %   \State \Return $S$, $F$
 \end{algorithmic}
\end{algorithm}

As shown in Algorithm~\ref{alg:umm}, UMM  starts with a \emph{small} sample of $\minit$ permutations that are randomly generated (although it would be possible to employ Latin Hyper-Cube sampling or other methods) and evaluated. Then, using the current sample $S$, the uBorda algorithm is applied to find a reference permutation $\hat{\sigma}_0$ (line~\ref{line:uborda}). Together with the parameter $\theta$, they define a Mallows model (Eq.~\ref{eq:MM}) from which we sample a new candidate permutation (line~\ref{line:sample}). This candidate permutation (solution) is evaluated and added to the sample $S$. The process is repeated up to a given maximum budget of evaluations $\FEmax$.\MANUEL{We do not explain the initial value of $\theta$ or how we decrease it. Can we give a bit more detail?}

@@

\subsubsection{CEGO.}
We use the implementation of GECO provided by the
authors.\footnote{\url{https://cran.r-project.org/package=CEGO}} %
Although it is never stated in the original paper, the implementation of CEGO
generates a set of initial solutions of size $\minit=10$ by means of a
max-min-distance sequential design: new solutions are added to the set
sequentially by maximizing the minimum distance to solutions already in the
set. These initial solutions are then evaluated on the true fitness function
and the result is used to build the initial surrogate
model.% \citet{ZaeStoBar2014:ppsn} uses \minit=10
Following the authors of
CEGO~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn}, we use a GA to optimize
the surrogate models with population size of 20, crossover rate of 0.5,
mutation rate of $1/n$, tournament selection of size 2 with probability of 0.9,
interchange mutation (i.e., exchanging two randomly selected elements) and
cycle crossover for permutations. To generate each new solution, CEGO runs the
GA using the model as a surrogate of the true fitness function.  The authors of
CEGO used a budget of $10^4$ surrogate-model evaluations for each run of the
GA. However, with such a setting, a single run of CEGO with a budget of $\FEmax=400$
true fitness evaluations requires \emph{more than a week} of computation time,
which was infeasible for our study. Hence, we set the budget of the GA to
$10^3$. Although a higher GA budget could improve the results of CEGO reported
here, our results are similar (and sometimes better) than the ones reported in
the original papers~\citep{ZaeStoFriFisNauBar2014,ZaeStoBar2014:ppsn} due to
running CEGO with $\FEmax=400$ instead of $\FEmax=200$. In fact, we found that the
increase in runtime caused by doubling $\FEmax$ is much smaller than by increasing
the GA budget by a factor of $10$. Nevertheless, we caution that choosing a
different trade-off may lead to different conclusions and more work is needed
to understand CEGO parameters.

\subsubsection{Other}.  For simplicity, we use Kendall's $\tau$
distance~(Eq.~\ref{eq:kendall_decomp}) in all
experiments\MANUEL{\citet{ZaeStoBar2014:ppsn} claims this is called Swap
  distance, could you double-check?}. Nevertheless,
\citet{ZaeStoFriFisNauBar2014} points out that CEGO performs best on the QAP
when using Hamming distance. We plan to extend our analysis of UMM to other
distance measures and to the dynamic selection of distance
measures~\citep{ZaeStoBar2014:ppsn}. xIn all experiments, we consider a maximum budget of $\FEmax=400$ evaluations of
the actual objective function. In a white-box context, state-of-the-art
algorithms for the problems considered here typically evaluate
\textcolor{red}{hundreds? thousands?} of solutions, thus, the budget considered
here for the black-box context is extremely limited.

@@

For the LOP, we generated synthetic instances of size $n=20$, $m=200$, $\phi\in\{0.5,0.7,0.9\}$\MANUEL{what is $m$ here? We already use $\FEmax$ for the budget, but we could change it. }
